---
title: "co2_DT"
author: "Ercong Luo"
date: "11/12/2021"
output: html_document
---

```{r}
library(tidyverse)
library(readxl)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(readr)
library(dplyr)
library(party)
library(ROCR)
set.seed(100)
```


```{r}
df <- read_excel("CO2_passenger_cars.xlsx", guess_max = 300, n_max = Inf)
```


Subseting the DF to relevant features: 

```{r}
features = c( "MS",
              "Mk",
              "Cn",
              "Ct",
              "Cr",
              "m (kg)",
              "Enedc (g/km)",
              "Ewltp (g/km)",
              "W (mm)",
              "Ft",
              "Fm",
              "ec (cm3)",
              "ep (KW)",
              "z (Wh/km)",
              "It",
              "Ernedc (g/km)",
              "Erwltp (g/km)",
              "De",
              "Vf",
              "r")
df = subset(df, select = features)
```


Metadata: 

```{r}
head(df, n = 30)
str(df)
```

Check which columns have too many _NA_ values: 

```{r}
for (feature in features){
  temp = mean(is.na(df[feature]))
  if (temp < 0.10){
    string = paste(feature,": proportion of NA values",  "is", temp)
  } else{
    string = paste("***", feature, ": proportion of NA values", "is TOO HIGH at", temp, "***")
  }
  print(string)
}
```

```{r}
df.model = subset(df, select = -c(Cr,
            `Ewltp (g/km)`,
            `z (Wh/km)`,
            It,
            `Ernedc (g/km)`,
            `Erwltp (g/km)`,
            De,
            Vf,
            r))
```

Cleaning up the make of the cars. And these are the electric car makers represented in the dataset:
```{r}
df$Mk = toupper(df$Mk)
df$Mk[df$Mk == "ALFA"] = "ALFA ROMEO"
df$Mk[df$Mk == "ASTON MARTIN"] = "ASTON-MARTIN"
df$Mk[df$Mk == "Audi"] = "AUDI"
df$Mk[df$Mk == "BMW ALPINA"] = "BMW"
df$Mk[df$Mk == "BMW I"] = "BMW"
df$Mk[df$Mk == "CITROËN"] = "CITROEN"
df$Mk[df$Mk == "CITROEN, DS"] = "CITROEN"
df$Mk[df$Mk == "Ford"] = "FORD"
df$Mk[df$Mk == "Honda"] = "HONDA"
df$Mk[df$Mk == "ALFA"] = "ALFAROMEO"
df$Mk[df$Mk == "MERCEDES"] = "MERCEDES-BENZ"
df$Mk[df$Mk == "TESLA MOTORS"] = "TESLA"
df$Mk[df$Mk == "VOLKSWAGEN, VW"] = "VOLKSWAGEN"
df$Mk[df$Mk == "VOLKSWAGEN,VW"] = "VOLKSWAGEN"
df$Mk[df$Mk == "VW"] = "VOLKSWAGEN"
(table(df$Mk[df$Ft == "ELECTRIC"]))
```

```{r}
df$Ft = toupper(df$Ft)
```

```{r}
df.model = df.model[(df.model$Ft== "DIESEL" | df.model$Ft== "ELECTRIC" | df.model$Ft== "PETROL"),]
boxplot(`Enedc (g/km)`~Ft, data = df.model, las=2)
```

```{r}
print(table(df.model$Mk,df.model$Ft))
```

```{r}
print(table(df.model$Ft))
```

More more round of data cleaning: convert to Factors when necessary. 

```{r}
df.model = subset(df.model, select = -Ct)
df.model$MS = factor(df.model$MS)
df.model$Mk = factor(df.model$Mk)
df.model$Ft = factor(df.model$Ft)
df.model$Fm = factor(df.model$Fm)
```

Split test-train datasets. 

```{r}
get.train.indices <- function(){
  index1 <- sample(df.model$tag[df.model$Ft == "ELECTRIC"], 50, replace = T)
  index2 <- sample(df.model$tag[df.model$Ft == "DIESEL"], 50, replace = T)
  index3 <- sample(df.model$tag[df.model$Ft == "PETROL"], 50, replace = T)
  return = c(index1, index2, index3)
}

df.model$tag = c(1:nrow(df.model))
temp = get.train.indices()
train = df.model[temp,]
test = df.model[-temp,]
```

Looking at the proportion of sampling, the proportions of each category in the test and training sets are comparable. 
```{r}
table(df.model$Ft)
table(train$Ft)#/nrow(data_train)
table(test$Ft)#/nrow(data_test)
```

```{r}
names(test)
```


Now train the tree models. Three trees will be created:

```{r}
hyperparams = rpart.control(maxdepth = 3)
tree = rpart(data = train, Ft ~ MS+Mk+`m (kg)`+`Enedc (g/km)`+`W (mm)`+`ec (cm3)`+`ep (KW)`, control = hyperparams)
(rpart.plot(tree))
pred = predict(tree, newdata = test, type="class")
(TABLE = table(pred, test$Ft))
```

```{r}
hyperparams = rpart.control(maxdepth = 3)
tree = rpart(data = train, Ft ~ MS+Mk+`m (kg)`+`W (mm)`+`ec (cm3)`+`ep (KW)`, control = hyperparams)
(rpart.plot(tree))
pred = predict(tree, newdata = test, type="class")
(TABLE = table(pred, test$Ft))
```

```{r}
hyperparams = rpart.control(maxdepth = 5)
tree = rpart(data = train, Ft ~ `m (kg)`+`W (mm)`+`ec (cm3)`+`ep (KW)`, control = hyperparams)
(rpart.plot(tree))
pred = predict(tree, newdata = test, type="class")
(TABLE = table(pred, test$Ft))
```




