---
title: "carEmissionData"
author: "Ercong Luo"
date: "9/29/2021"
output: html_document
---

```{r}
library(readxl) # to read in excel sheet
```


First to read the excel sheet in as a dataframe:
```{r}
DF <- read_excel("CO2Data_raw.xlsx",)
(str(DF))
```

As seen below, the tail of the dataset contains two rows of data that are only footnotes. So they 
will be deleted.
```{r}
(tail(DF))
DF <- DF[-c(22,23),]
(tail(DF))
```

Next, fix data types":\

- "Model" is good as is. 
- "Make" should be of type _factor_. 
- "EV" should be of type _logical_. 
- "Model Year" should be of type _integer_. 
- "Units Sold" should be of type _integer_, they aren't long enough to be doubles. 
- "Battery Emissions" should be of type _numeric_
- "Engine/Motor Emissions (tons)", "Maintenance Emissions (tons)", "Body Construction (tons)", "Tank Emissions (tons)" should be of type _numeric_
- "EV Feature" should be of type _factor_

To the following chunk of code will fix the data types
```{r}
DF$Make <- as.factor(DF$Make)
DF$EV<- as.logical(DF$EV)
DF$`Model Year`<- as.integer(DF$`Model Year`)
DF$`Units sold` <- as.integer(DF$`Units sold`)
DF$`Battery Emissions (tons)` <- as.numeric(DF$`Battery Emissions (tons)`)
DF$`Engine/Motor Emissions (tons)` <- as.numeric(DF$`Engine/Motor Emissions (tons)`)
DF$`Maintenance Emissions (tons)` <- as.numeric(DF$`Maintenance Emissions (tons)`)
DF$`Body Construction (tons)` <- as.numeric(DF$`Body Construction (tons)`)
DF$`Tank Emissions (tons)` <- as.numeric(DF$`Tank Emissions (tons)`)

# the NA values in EV Feature stand for "not applicable" because this feature only applies to EVs. 
DF$`EV Feature`[DF$EV == FALSE] = "NotApplicable"
DF$`EV Feature` <- as.factor(DF$`EV Feature`)
```

```{r}
str(DF)
```

```{r}
DF
```

It can be seen above that all Battery Emission entries with NA values are internal-combustion cars, so obviously there was so
CO2 emissions with producing its battery pack. Therefore these entries can just be filled in with zeros:
```{r}
DF$`Battery Emissions (tons)`[DF$EV == FALSE] = 0
```
The emissions for producing an engine for internal combustion cars has a pretty small standard deviation (0.217 as calculated below). Therefore it is 
okay to replace NA values with the mean of internal combustion cars. 
```{r}
(sd(DF$`Engine/Motor Emissions (tons)`[!is.na(DF$`Engine/Motor Emissions (tons)`) & DF$EV == FALSE]))
temp <- (mean(DF$`Engine/Motor Emissions (tons)`[!is.na(DF$`Engine/Motor Emissions (tons)`) & DF$EV == FALSE]))
DF$`Engine/Motor Emissions (tons)`[is.na(DF$`Engine/Motor Emissions (tons)`) & DF$EV == FALSE] <- temp
```

Other features where the variation of the data is small and it is okay to plug in mean value is the Tank Emissions and Body Construction, and Maintenance Emissions.
```{r}
(sd(DF$`Tank Emissions (tons)`[!is.na(DF$`Tank Emissions (tons)`) & DF$EV == FALSE]))
temp <- (mean(DF$`Tank Emissions (tons)`[!is.na(DF$`Tank Emissions (tons)`) & DF$EV == FALSE]))
DF$`Tank Emissions (tons)`[is.na(DF$`Tank Emissions (tons)`) & DF$EV == FALSE] <- temp
```

```{r}
(sd(DF$`Body Construction (tons)`[!is.na(DF$`Body Construction (tons)`)]))
temp <- (mean(DF$`Body Construction (tons)`[!is.na(DF$`Body Construction (tons)`)]))
DF$`Body Construction (tons)`[is.na(DF$`Body Construction (tons)`)] <- temp
```

```{r}
(sd(DF$`Maintenance Emissions (tons)`[!is.na(DF$`Maintenance Emissions (tons)`) & DF$EV == FALSE]))
temp <- (mean(DF$`Maintenance Emissions (tons)`[!is.na(DF$`Maintenance Emissions (tons)`) & DF$EV == FALSE]))
DF$`Maintenance Emissions (tons)`[is.na(DF$`Maintenance Emissions (tons)`) & DF$EV == FALSE] <- temp
```

Now that the record data is cleaned, write this out to a .csv file. 
```{r}
write.csv(DF, file = "carEmissionData_cleaned.csv")
```

